Generalized measurement invariance tests with application to factor analysis

### 2011-11-17 www/paper.Rnw
further small touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@78 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Method}
Wicherts et al.\ tested a series of confirmatory factor models for
various types of measurement invariance.  We focus on the model used
in their Step 5b (see pp.\ 703--704 of their paper), which involved
across-group parameter restrictions on the factor loadings, unique
variances, intercepts, and factor variances.  These parameters were
typically restricted to be equal across groups, though a subset of the
parameters were allowed to be group-specific upon examination of
modification indices.  Most notably, the most difficult intelligence test's factor loading, unique variance, and intercept differed for the ``minority,
stereotype threat'' group, as compared to the other groups.
Factor means were also estimated for three of the groups.  The model
provided a reasonable fit to the data, as judged by examination of
$\chi^2$, RMSEA, and CFI statistics.  Wicherts et al.\ also had a Step
6 (the final step), where they examined a model that was simpler than
described above and that provided a slightly better fit to the data.
This model restricted factor means to be equal across a subset of the
groups, however, which was not helpful for our purposes (because we are focusing on a single group).

@@

Test statistics for the hypotheses \eqref{eq:exh0}
and~\eqref{eq:exh0.2} are displayed in Figure~\ref{fig:cusumex}.
Each panel displays a test 
statistic's fluctuation across values of student GPA, with the first
column containing tests of~\eqref{eq:exh0} (testing $k*=4$ parameters)
and the second column containing tests of~\eqref{eq:exh0.2} (testing
$k*=1$ parameter).
The solid
horizontal 
lines represent critical values for $\alpha= 0.05$, and
the Cram\'{e}r-von Mises panels 
also contain a dashed line depicting the value of the test
statistic (test statistics for the others are simply 
the maxima of the processes).
In other words, for panels in the first and third rows, \eqref{eq:exh0} is
rejected if the process crosses the horizontal line.  For panels in
the second row, \eqref{eq:h0} is rejected if the dashed horizontal line is
higher than the solid horizontal line.

%% Details on test results
The results displayed in Figure~\ref{fig:cusumex} generally show that the
``minority, stereotype prime'' group lacks invariance with respect to
GPA.  All three tests are significant at the .05 level when the four
group-specific parameters are tested simultaneously (first column).
When the intercept parameter is tested individually, two of the three
tests achieve significance (the CvM test and $\max \mathit{LM}$
statistics).  These results generally support the notion that domain
identification (as measured by GPA) influences the magnitude of
stereotype threat.  The $k*=1$ results also imply that the tests
differ in power and/or in error rates, an issue we address later in the
paper.

%% add paragraph about peaks of graphs
We can also use the test results to estimate the threshold $\nu$,
which in this context can be used to define GPA-based subgroups whose
measurement parameters differ.  As described previously, estimates of
$\nu$ can be obtained by 
examining the peaks in Figure~\ref{fig:cusumex}.  For the $k*=4$
panels (first column), two peaks occur near GPAs of 6.0 and of 6.3 (6
corresponds to the American `C' range).  These peaks roughly divide
individuals into those receiving D's and F's, those receiving solid
C's, and those receiving high C's, B's, and A's.  Parameter-specific
results can also be obtained, as shown for the intercept parameter in
the second column of Figure~\ref{fig:cusumex}.  Here, a single peak
occurs in the three panels around a GPA of 7.0 (`B' range).

>>>>

\subsection{Method}
Wicherts et al.\ tested a series of confirmatory factor models for
various types of measurement invariance.  We focus on the model used
in their Step~5b (see pp.~703--704 of their paper), which involved
across-group parameter restrictions on the factor loadings, unique
variances, intercepts, and factor variances.  These parameters were
typically restricted to be equal across groups, though a subset of the
parameters were allowed to be group-specific upon examination of
modification indices.  Most notably, the most difficult intelligence test's factor loading, unique variance, and intercept differed for the ``minority,
stereotype threat'' group, as compared to the other groups.
Factor means were also estimated for three of the groups.  The model
provided a reasonable fit to the data, as judged by examination of
$\chi^2$, RMSEA (root-mean-square error of approximation), and CFI
(comparative fit index) statistics.  Wicherts et al.\ also had a
Step~6 (the final step), where they examined a model that was simpler than
described above and that provided a slightly better fit to the data.
This model restricted factor means to be equal across a subset of the
groups, however, which was not helpful for our purposes (because we are focusing on a single group).

@@

Test statistics for the hypotheses \eqref{eq:exh0}
and~\eqref{eq:exh0.2} are displayed in Figure~\ref{fig:cusumex}.
Each panel displays a test 
statistic's fluctuation across values of student GPA, with the first
column containing tests of~\eqref{eq:exh0} (testing $k^*=4$ parameters)
and the second column containing tests of~\eqref{eq:exh0.2} (testing
$k^*=1$ parameter).
The solid
horizontal 
lines represent critical values for $\alpha= 0.05$, and
the Cram\'{e}r-von Mises panels 
also contain a dashed line depicting the value of the test
statistic (test statistics for the others are simply 
the maxima of the processes).
In other words, for panels in the first and third rows, \eqref{eq:exh0} is
rejected if the process crosses the horizontal line.  For panels in
the second row, \eqref{eq:h0} is rejected if the dashed horizontal line is
higher than the solid horizontal line.

%% Details on test results
The results displayed in Figure~\ref{fig:cusumex} generally show that the
``minority, stereotype prime'' group lacks invariance with respect to
GPA.  All three tests are significant at the .05 level when the four
group-specific parameters are tested simultaneously (first column).
When the intercept parameter is tested individually, two of the three
tests achieve significance (the CvM test and $\max \mathit{LM}$
statistics).  These results generally support the notion that domain
identification (as measured by GPA) influences the magnitude of
stereotype threat.  The $k^*=1$ results also imply that the tests
differ in power and/or in error rates, an issue we address later in the
paper.

%% add paragraph about peaks of graphs
We can also use the test results to estimate the threshold $\nu$,
which in this context can be used to define GPA-based subgroups whose
measurement parameters differ.  As described previously, estimates of
$\nu$ can be obtained by 
examining the peaks in Figure~\ref{fig:cusumex}.  For the $k^*=4$
panels (first column), two peaks occur near GPAs of 6.0 and of 6.3 (6
corresponds to the American `C' range).  These peaks roughly divide
individuals into those receiving D's and F's, those receiving solid
C's, and those receiving high C's, B's, and A's.  Parameter-specific
results can also be obtained, as shown for the intercept parameter in
the second column of Figure~\ref{fig:cusumex}.  Here, a single peak
occurs in the three panels around a GPA of 7.0 (`B' range).
